\section{Conclusion}
\label{sec:conclusion}

We presented a conformance-gated benchmark methodology and applied it to
what is, to our knowledge, the first empirical comparison of four
digit-generation algorithms (\burgerdybvig, multiprecision; \schubfach,
\ryualg, and \dragonbox, all fixed-width) inside a
\rfcjcs/\ecma-constrained JSON canonicalization pipeline, with
cross-language validation in Go and Rust to separate algorithmic effects
from language/runtime characteristics.  While prior work has compared
float-to-string algorithms in isolation (e.g., Adams's comparison of
Ry\=u against Grisu and Errol; Champagne Gareau and Lemire's
comprehensive C++ evaluation), ours is, to our knowledge, the first
study to evaluate these algorithms within a specification-constrained
canonicalization pipeline using conformance-gated methodology.

Key findings:
\begin{itemize}
\item The three fixed-width algorithms (\schubfach, \ryualg, \dragonbox)
      perform within 1--8\% of each other across both languages and all
      workloads.  The dominant performance axis is fixed-width vs.\
      multiprecision, not the choice among fixed-width algorithms.
\item Fixed-width algorithms achieve 1.4--1.8$\times$ speedup on
      number-dense workloads ($\nu(D) \geq 40\%$) in Go API benchmarks
      due to zero heap allocation; the advantage replicates at
      2.3--3.0$\times$ in Rust (excluding extreme outliers); the
      replication of within-language ordering across both languages
      provides evidence that the advantage is not solely attributable
      to language-specific runtime characteristics.
\item In end-to-end CLI measurement, the advantage narrows due to
      process startup dominance, but remains statistically significant
      on number-dense workloads.
\item The advantage narrows to $<$5\% on string-dominant and
      deeply-nested inputs, correlating with number density $\nu(D)$.
\item Go implementations produce byte-identical output under
      cross-compilation for x86-64 and arm64 (arm64 binaries verified
      via QEMU user-mode emulation on x86-64 host).  Rust
      implementations were not cross-compiled in this study.
\item Preliminary profiling suggests digit generation accounts for
      approximately 15--40\% of total canonicalization cost depending
      on number density. The observed pipeline-level speedups
      (1.4--1.8$\times$ in Go, 2.3--3.0$\times$ in Rust) are consistent
      with isolation-level speedups of ${\approx}5\times$ attenuated by
      Amdahl's Law at these pipeline fractions.
\end{itemize}

The benchmark infrastructure, 286{,}362 oracle vectors, and all raw data
are released as a reproducibility artifact.  Our findings align with the
broader ecosystem direction: Go~1.26 adopted the Dragonbox fixed-width
algorithm for \texttt{strconv}, corroborating our evidence that
fixed-width approaches offer meaningful performance advantages in
float-to-string formatting.  Since the three fixed-width algorithms are
performance-equivalent, implementors can choose based on other criteria:
code simplicity, ecosystem momentum, or theoretical interest.

\subsection*{Future Work}

Bare-metal arm64 performance benchmarks, additional language
implementations (C and JVM) to extend the comparison matrix to unmanaged
and managed-GC runtimes, Rust cross-compilation verification, and
workload characterization from production JSON corpora.
