\section{Discussion}
\label{sec:discussion}

\subsection{Fixed-Width Algorithm Equivalence}

The most striking finding is that the three fixed-width algorithms
(\schubfach, \ryualg, \dragonbox) perform within 1--8\% of each other
across both languages.  In Go API benchmarks, pairwise comparisons among
fixed-width algorithms yield speedups of ${\leq}1.02\times$, with only
6~of~89 go:ryu-vs-dragonbox comparisons reaching BH significance.
In Rust, the pattern is similar: 9--12 BH-significant comparisons
out of 76 per pair, with speedups of ${\leq}1.13\times$ and no
consistent winner.

This equivalence has a mechanistic explanation: all three algorithms
share the same computational core---128-bit multiply against a
precomputed power-of-ten table, followed by digit trimming.  The
differences (\ryualg's lastRemovedDigit heuristic, \dragonbox's
two-digit-at-a-time extraction, \schubfach's bound computation) affect
branch structure but not the dominant cost: the 64$\times$128-bit
multiplication and table lookup.  On modern x86-64 hardware with fast
\texttt{MUL} throughput, the branch differences are below measurement
noise in the pipeline context where parsing and serialization dominate.

\subsection{Workload-Dependent Behavior}

The performance advantage of fixed-width algorithms over \burgerdybvig is
strongly workload-dependent.  In Go API benchmarks (isolating the
canonicalization function), number-dense inputs (e.g.,
\texttt{array-2048}, \texttt{large}, \texttt{medium}) show
1.4--1.8$\times$ speedup due to zero heap allocation in the
digit-generation path.  In Rust within-language API benchmarks, the same
algorithmic ordering holds at larger magnitudes (2.3--3.0$\times$ on the
same workloads, excluding the two extreme outliers discussed below),
providing evidence that the advantage is not solely attributable to
language-specific runtime characteristics.  Deeply-nested
workloads (\texttt{deep}, \texttt{deep-64}) contain numbers but show
${\approx}1.0\times$ because JSON structural parsing and key sorting
dominate, diluting the digit-generation signal.  The speedup therefore
correlates with number \emph{density} (fraction of total cost attributable
to digit generation), not merely number presence.
Figure~\ref{fig:density-speedup} visualizes this relationship: workloads
with zero number content cluster near 1.0$\times$, while number-dense
workloads ($>$70\% number values) show the full algorithmic advantage.

\input{figures/number-density-speedup}

\subsection{Process Startup Dominance}

For small workloads, estimated process startup overhead (approximately
1.3\,ms) accounts for roughly 87--93\% of CLI execution time (e.g.,
\texttt{canonical-minimal} total $\approx$1.47\,ms in canonicalize
mode).  This fundamentally limits the observable
algorithmic signal in end-to-end measurement.  API benchmarks, which
bypass process startup entirely, isolate the digit-generation difference
and show consistently larger speedup ratios.  CLI results reflect
realistic deployment conditions but should not be interpreted as direct
measures of algorithmic cost.

\subsection{Allocation Cost Attribution}

The \burgerdybvig implementation's \texttt{sync.Pool} partially mitigates
allocation cost by reusing big-integer buffers.  Without pooling, the
difference would be larger.  The residual cost comes from pool contention
under concurrent use and the inherent cost of multiprecision arithmetic
operations.

API benchmark allocation data directly attributes the speedup:
on \texttt{array-2048}, fixed-width algorithms (\schubfach, \ryualg,
\dragonbox) perform ${\approx}$49{,}134 allocs/op vs.\
json-canon's 57{,}343 allocs/op ($\Delta$4.0 per number, 91\,B per
number).  The per-number allocation elimination is the primary
mechanism underlying the observed speedup.

\subsection{Extreme Outliers: \texttt{numeric-boundary} and \texttt{number-heavy}}

\input{figures/outlier-speedup}

Two workloads show unusually large speedup ratios in Rust API
benchmarks: \texttt{numeric-boundary} (${\approx}9.3\times$ within-Rust,
CI [9.1, 9.5]) and \texttt{number-heavy} (${\approx}8.2\times$,
CI [8.1, 8.4]), compared to ${\approx}1.5\times$ in Go for both.
All three fixed-width algorithms show similar outlier ratios against
\burgerdybvig on these workloads, further confirming that the effect
is a fixed-width vs.\ multiprecision phenomenon.
Because these are \emph{within-Rust} comparisons (same harness on both
sides), they are unaffected by the cross-language measurement granularity
asymmetry discussed in \S\ref{sec:methodology}.  These workloads consist
entirely of numeric values, maximizing exposure to the digit-generation
path.  We report these workloads separately because their near-zero
non-numeric content makes the pipeline-level ratio approach the
isolation-level ratio, which is not representative of mixed workloads
where parsing and serialization dilute the digit-generation signal.  The Rust \burgerdybvig implementation uses the \texttt{num-bigint}
crate (Karatsuba multiplication) with a lazily-initialized
power-of-ten cache (\texttt{OnceLock<Vec<BigUint>>}); although the
harness discards 3~warmup iterations to amortize first-invocation
costs, the cache-warmed steady-state cost of
multiprecision arithmetic remains
substantially higher than the fixed-width implementations'
compile-time constant lookup table.  The tight confidence intervals
and low MDEs (2--3\%) confirm these are genuine algorithmic
cost differences, not measurement noise.  The Go within-language
comparisons at ${\approx}1.5\times$ are smaller because Go's
\texttt{sync.Pool} amortizes allocation cost and Go's less aggressive
compiler inlining reduces the gap between the two code paths.

Table~\ref{tab:per-iteration-outliers} shows per-iteration timing
stability for these outlier workloads.  Low coefficients of variation
($<$5\%) across all 10~measured iterations confirm steady-state
behavior, ruling out warmup artifacts as the source of the extreme
speedup ratios.

\input{tables/per-iteration-outliers}

\subsection{Profile-Level Analysis}

Preliminary CPU profiling (via Go's \texttt{pprof}) suggests that digit
generation accounts for approximately 15--40\% of total canonicalization
time depending on workload number density.  On \texttt{array-2048}
($\nu = 40\%$), digit generation accounts for $\approx$35\% of CPU time
in the \burgerdybvig path; on \texttt{deep} ($\nu = 0\%$), it drops to
$<$2\%.  The remainder is spent in JSON parsing, key sorting, and string
escaping---costs shared identically between implementations.  Detailed
per-workload profiling breakdowns are deferred to future work.

\subsection{Practical Implications}

For applications where canonicalization throughput is critical (e.g.,
high-frequency signature verification of number-heavy payloads),
any of the three fixed-width algorithms offers meaningful improvement
over \burgerdybvig.  On the \texttt{medium} workload (13.7\,KB,
$\nu = 40\%$), the Go API difference between fixed-width algorithms
(${\approx}342\,\mu$s/op) and \burgerdybvig
(${\approx}581\,\mu$s/op) is approximately 239\,$\mu$s per
canonicalization.  A service verifying 1{,}000 signed JSON
documents per second at this payload size would reclaim roughly
0.24~seconds of CPU time per second---freeing approximately 24\% of
a core compared to the \burgerdybvig path.

The engineering context that produced these results is illustrative.
The \burgerdybvig implementation was replaced with \schubfach in a
production canonicalization library, and the predicted improvement
materialized.  The question of whether the choice \emph{among}
fixed-width algorithms mattered then motivated this study.  It does
not: all three are equivalent within 1--8\%.  Implementors can
therefore choose based on code simplicity (\schubfach has the smallest
implementation), ecosystem momentum (\dragonbox was adopted by
Go~1.26's \texttt{strconv}), or historical priority (\ryualg as the
first full-coverage fixed-width algorithm), confident they are not
leaving performance on the table.

\subsection{Cross-Language Interpretation}

Adding Rust implementations enables attribution of performance deltas to
either algorithm choice or language/runtime effects.  The within-language
comparisons test whether the fixed-width advantage persists independent of
Go's allocator and garbage collector behavior, while same-algorithm
cross-language comparisons quantify language overhead (startup and
steady-state) directly.

Rust's ownership model removes the need for Go-style
\texttt{sync.Pool}-based reuse patterns.  For multiprecision paths, this
shifts lifecycle cost from GC-mediated reuse to deterministic RAII
destruction, which can materially change allocation profiles and latency
variance characteristics.

\textbf{Startup overhead.}  Rust native binaries exhibit substantially
lower process startup cost than Go binaries (roughly 2--3$\times$ faster
in CLI wall time), because they avoid Go runtime initialization (scheduler
bootstrap, GC metadata setup, goroutine stack allocation).  This
amplifies the startup-dominance effect: for small workloads the language
runtime overhead is the primary differentiator, not the digit-generation
algorithm.

\textbf{Algorithmic ordering replicates across languages.}  The key finding
from the cross-language design is that within-Rust comparisons replicate
the within-Go ordering: all three fixed-width algorithms outperform
\burgerdybvig on number-dense workloads in both languages, and the three
fixed-width algorithms are equivalent in both languages.  This provides
evidence that the performance advantage derives from the fixed-width
approach itself (zero heap allocation, no multiprecision arithmetic),
not from language-specific runtime characteristics.

\textbf{Interaction analysis.}  The eight implementations form
a $4 \times 2$ structure (algorithm $\times$ language).  While we do not
perform a formal factorial interaction test due to measurement granularity
asymmetry between harnesses, Table~\ref{tab:interaction-analysis} reports
a ratio-of-ratios (RoR) analysis for the \schubfach vs.\ \burgerdybvig
pair: for each canonicalize workload, $R_\text{Go}$ and $R_\text{Rust}$
are the within-language speedup ratios, and
$\text{RoR} = R_\text{Rust}/R_\text{Go}$.

\input{tables/interaction-analysis}

$\text{RoR} > 1$ indicates that the fixed-width advantage is amplified in
Rust.  On number-dense workloads, $\text{RoR}$ consistently exceeds 1
(typically 1.3--2.0$\times$), reflecting the larger allocation cost
differential in Rust where the \texttt{num-bigint}-based path lacks Go's
\texttt{sync.Pool} amortization.  Readers should interpret the RoR as
descriptive evidence of interaction magnitude, not a formal statistical
test of interaction significance.

\textbf{Cross-language comparison caveats.}  Same-algorithm
cross-language API comparisons (e.g., schubfach:go-vs-rs) should be
interpreted carefully.  Go's \texttt{testing.B} averages over many
inner iterations per sample, producing low-variance measurements, while
the Rust harness times individual function calls, producing
higher-variance samples.  Additionally, the two compilers apply
fundamentally different optimization strategies: Rust with LTO and
\texttt{codegen-units=1} enables aggressive cross-crate inlining that
Go's compiler cannot match.  The observed 2--3$\times$ cross-language
speedup (API) may partly reflect compiler optimization quality rather
than inherent language characteristics.  These comparisons primarily
serve to confirm directional consistency of the algorithmic ordering,
not to quantify precise language overhead.

\textbf{Allocation instrumentation.}  The Rust bench harness uses a
counting global allocator (\texttt{std::alloc::GlobalAlloc}) that
increments atomic counters on each allocation, reporting per-operation
byte and allocation counts in the same format as Go's
\texttt{testing.B}.  This confirms the mechanistic explanation:
Rust fixed-width implementations allocate minimally (e.g., 1{,}036~B/op,
24~allocs/op on \texttt{numeric-boundary}) while Rust-\burgerdybvig
allocates substantially more (25{,}204~B/op, 470~allocs/op on the same
workload) due to \texttt{num-bigint} multiprecision arithmetic.  The
$24.3\times$ allocation byte ratio directly explains the
${\approx}9\times$ speedup: allocation and initialization cost dominates
the multiprecision path.

\subsection{Comparison with Isolation Benchmarks}

Champagne Gareau and Lemire~\cite{champagnegareau2026} report
substantially larger speedup ratios for fixed-width algorithms over
Dragon4 in C++ isolation benchmarks, where only the digit-generation
function is timed.  The difference from our pipeline ratios is
explained by Amdahl's Law: if digit generation accounts for fraction
$P$ of total pipeline cost and provides speedup $S$ in isolation,
the pipeline-level speedup is $1/(1 - P + P/S)$.  Our profiling data
(\S\ref{sec:discussion}, Profile-Level Analysis) shows $P \approx 0.35$
on number-dense workloads.  With a representative isolation speedup of
$S = 5\times$, Amdahl's Law predicts a pipeline speedup of
$1/(0.65 + 0.07) = 1.39\times$, consistent with our observed
1.4--1.8$\times$ Go API ratios.  The remaining gap between prediction
and observation reflects differences in allocation overhead (our
pipeline includes JSON parsing and serialization, which share the
allocator with digit generation) and cache effects absent from
isolation benchmarks.  Moreover, the pipeline-level speedup on these
workloads is bounded by $1/(1-P) \approx 1.54\times$---the Amdahl
ceiling when $P \approx 0.35$.  The observed Go API ratios of
1.4--1.8$\times$ already approach this ceiling, indicating that the
fixed-width algorithms have largely exhausted the available improvement
from optimizing digit generation alone; further gains require reducing
parsing or serialization cost.  This convergence confirms that our
pipeline results and their isolation results measure the same underlying
algorithmic advantage at different levels of system integration.

\subsection{Methodological Contribution}

The conformance-gated methodology introduced in this paper is
independent of the specific algorithms compared.  The \rfcjcs
canonicalization pipeline serves as the case study in which the
methodology was developed and validated; the methodology itself is
general, applicable to any specification-constrained formatting
evaluation where correctness is a binary prerequisite.  The core
design---gate timing on correctness, validate against multiple
independent oracles, document noise floors and minimum detectable
effects for every comparison---was not adopted as an academic
principle but developed from the practical discovery that the standard
ecosystem pattern (delegating number formatting to a language runtime)
produces implementations that pass casual testing while failing
specification-level verification.  The methodology addresses a gap in
the float-to-string literature, where prior benchmarks (including
Adams's Ry\=u evaluation and Champagne Gareau and Lemire's
comprehensive C++ study) measure performance in isolation without
first establishing conformance to a target specification.  In domains
where correctness is a hard prerequisite---canonicalization,
serialization standards, interoperability testing---measuring the
speed of a non-conformant implementation is meaningless.

As a concrete example of generalizability, consider CBOR deterministic
encoding (RFC~8949 \S4.2)~\cite{rfc8949}: the specification requires a
specific integer-to-bytes mapping and canonical map key ordering.  The
conformance-gated approach applies directly---oracle vectors would be
generated from a reference CBOR encoder, and candidate implementations
would be gated on byte-identical output before timing comparison of
encoding strategies.  The benchmark infrastructure released with this
paper implements this methodology as a reusable pipeline.
