\section{Implementation}
\label{sec:implementation}

We maintain eight implementations with a common CLI ABI:
\begin{itemize}
\item \texttt{impl-schubfach} (Go): \schubfach number formatting.
\item \texttt{impl-json-canon} (Go): \burgerdybvig number formatting.
\item \texttt{impl-ryu} (Go): \ryualg number formatting.
\item \texttt{impl-dragonbox} (Go): \dragonbox number formatting.
\item \texttt{impl-schubfach-rs} (Rust): \schubfach-aligned Rust path.
\item \texttt{impl-json-canon-rs} (Rust): \burgerdybvig-aligned Rust path.
\item \texttt{impl-ryu-rs} (Rust): \ryualg-aligned Rust path.
\item \texttt{impl-dragonbox-rs} (Rust): \dragonbox-aligned Rust path.
\end{itemize}

All eight implementations contain their own digit-generation code and do
not delegate to the standard library's float-to-string routines (e.g.,
Go's \texttt{strconv.FormatFloat} or Rust's \texttt{std::fmt}).  This is
necessary because the standard routines do not implement \ecma formatting
policy.  Consequently, Go~1.26's adoption of Dragonbox for
\texttt{strconv}~\cite{godragonboxcl} does not affect our measurements.

Within each language, all four implementations share identical JSON
parser, serializer, error taxonomy, and \ecma formatting code; only the
\texttt{jcsfloat} digit-generation module differs.\footnote{A
\texttt{c0}-tracking correction in the \dragonbox two-digit trimming
loop (\texttt{removed~/~10} $\to$ \texttt{removed~\%~10}) was
identified during a pre-submission audit and corrected in both the Go
and Rust implementations.  A related variable-ordering fix was applied
to the three Rust fixed-width implementations.  Both corrections are
correctness-neutral for IEEE~754 inputs (all 286{,}362 oracle vectors
pass identically before and after the fix) but restore logical
consistency with the algorithm specification.}  This isolation
ensures that performance differences are attributable to the
digit-generation algorithm alone.

The lab harness treats each implementation as a black-box canonicalizer
with identical CLI contracts (\texttt{canonicalize}, \texttt{verify},
exit-code policy, and error-class expectations).  This preserves
algorithm comparability while allowing language/runtime differences to be
measured independently.

Rust builds use release configuration (\texttt{opt-level=3},
\texttt{lto=true}, \texttt{codegen-units=1}, \texttt{strip=true}) to
align optimization policy with production benchmarking practice.

\subsection{Rust Implementation Notes}

All four Rust implementations use native \texttt{u128} arithmetic
for the 128-bit intermediate products required by the fixed-width
algorithms, eliminating the need for compiler intrinsics or multi-word
emulation.
The Rust \burgerdybvig implementation uses the \texttt{num-bigint}
crate (v0.4), which provides Karatsuba multiplication for large
operands, comparable in optimization level to Go's
\texttt{math/big.Int} with its assembly-accelerated x86-64 inner
loops.  Buffer lifecycle is managed by Rust's RAII ownership model
(deterministic \texttt{Drop}), removing the need for explicit pooling
strategies such as Go's \texttt{sync.Pool}.

All Rust implementations share an identical JSON parser and serializer;
only the \texttt{jcsfloat} digit-generation module differs between them,
mirroring the Go design.  Rust binaries are approximately 424--456\,KB
(stripped), compared to 2.5--2.7\,MB for the statically-linked Go
binaries.

\subsection{Platform Independence and Fused-Multiply-Add}

IEEE~754 compliance guarantees that each floating-point result is
correctly rounded to destination precision, but it does not guarantee
that the same source expression produces identical bits on all
platforms.  Fused-multiply-add (FMA) instructions compute $a \times b +
c$ with a single rounding rather than two, producing results that can
differ by one ULP from the unfused sequence.  Go's compiler permits FMA
for any multiply-add expression~\cite{go17895}; the opt-out is an
explicit \texttt{float64()} cast forcing intermediate rounding.  A
real-world consequence was observed in December 2025, when a Go database
(DoltHub) produced different query plans on ARM vs.\ x86 because the
compiler emitted FMA on ARM but not x86, causing a one-ULP difference
that flipped a cost comparison~\cite{dolthub2025fma}.

For digit-generation algorithms used in canonicalization, where the
formatted bytes are cryptographically bound, platform-dependent output
is unacceptable.  The \burgerdybvig implementation with
\texttt{math/big.Int} is immune by construction: all operations execute
on the CPU's integer ALU, where floating-point fusion is inapplicable by
type.  The three fixed-width algorithms (\schubfach, \ryualg,
\dragonbox) use \texttt{math/bits.Mul64} and \texttt{u128} arithmetic
for their core computations; we verified that no floating-point
expressions in the digit-generation path are fusible.  All eight
implementations therefore satisfy the integer-arithmetic-only property
required for cross-platform determinism.
