\section{Introduction}
\label{sec:introduction}

Specification-constrained formatting is common in systems where
serialized bytes are hashed, signed, or used as consensus inputs:
a single divergent byte invalidates every downstream hash and
signature.  Yet the standard benchmarking approach for formatting
code, which measures speed in isolation and assumes correctness, is
inadequate here.  A fast implementation that silently deviates from
the specification is unusable, because the
deviation may not surface until production signatures fail to verify.
This paper introduces \emph{conformance-gated benchmarking}, a
methodology that treats specification correctness as a binary
prerequisite: candidates must pass an oracle vector suite before any
timing data is collected.  The methodology documents noise floors and
minimum detectable effect sizes for every comparison, applies
Benjamini--Hochberg FDR correction across the full comparison matrix,
and is general, applicable to any domain where correctness is a hard
prerequisite for meaningful performance measurement.

The methodology was developed from an engineering problem, not adopted
as an academic principle.  Building a correct JSON Canonicalization
Scheme (\rfcjcs)~\cite{rfc8785} primitive required implementing
\ecma \texttt{Number::toString}~\cite{ecma262} number formatting from
scratch, because existing \rfcjcs implementations delegate to language
runtimes (Go's \texttt{strconv.FormatFloat}, for example) that do
not implement the \ecma formatting policy (they make different notation
choices at key boundaries: \texttt{1e20} vs.\ the required
\texttt{100000000000000000000}; \texttt{1e-06} vs.\ the required
\texttt{0.000001}) and whose underlying algorithms have changed across
versions.  Conformance verification against 286{,}362 oracle vectors
generated from V8 and independently validated against SpiderMonkey and
JavaScriptCore confirmed these gaps.  A correct \burgerdybvig~\cite{burger1996printing}
implementation with exact multiprecision arithmetic was built, validated
against the full vector suite, and deployed; profiling then revealed
its heap allocation as a bottleneck on number-dense workloads.
Replacing \burgerdybvig with the fixed-width
\schubfach~\cite{giulietti2022schubfach} algorithm eliminated
digit-generation allocation entirely while producing byte-identical
output.  The natural follow-up question, \emph{is \schubfach
specifically the right choice, or would \ryualg~\cite{adams2018ryu} or
\dragonbox~\cite{jeon2021dragonbox} perform equivalently?}, asked with
statistical rigor inside a pipeline context, is the case study of this
paper.

Several deployment contexts place canonicalization on the critical path.
In high-frequency signature verification, every document must be
canonicalized before its signature can be checked, making formatting
throughput a direct bottleneck.  Content-addressed storage systems hash
the canonical form at both ingestion and retrieval, so canonicalization
cost scales linearly with object throughput.  Distributed consensus
protocols require byte-identical serialization for state-transition
agreement, invoking canonicalization on every proposal and vote.  In
each scenario, a constant-factor improvement in the formatting stage
translates directly to throughput gains: the 1.4--1.8$\times$ speedup
we measure on number-dense payloads corresponds to freeing
approximately 24\% of a CPU core at 1{,}000 canonicalizations per
second on a representative 14\,KB workload.

Beyond throughput, the choice of digit-generation algorithm is a
correctness concern in systems where canonicalized bytes are hashed and
signed at every layer of a processing pipeline.  In deployment contexts
such as deterministic infrastructure, where data-processing steps
execute in sandboxed isolation and every governed artifact is
schema-validated, canonically serialized, and cryptographically
bound, a formatting divergence at the number-rendering layer does not
produce one wrong document: it
can invalidate hashes computed over those bytes, signatures
computed over those hashes, and audit trails that chain those
signatures.  The digit-generation algorithm sits at the foundation of a
determinism guarantee that higher layers cannot independently recover.
This makes both the correctness validation (conformance gating) and the
performance characterization reported here directly relevant to
infrastructure where canonicalization is not optional but
load-bearing~\cite{rfc8785}.

This paper makes three contributions:

\begin{enumerate}
\item A conformance-gated benchmark methodology, developed from the
      practical experience that delegating to standard library formatters
      produces output that passes casual testing but fails
      specification-level verification.  The methodology gates
      performance measurement on correctness: implementations must pass
      286{,}362 oracle vectors and 776 structural conformance cases
      before timing data is collected.  We document noise floors and
      minimum observable effect sizes for all comparisons.  The
      methodology is general and reusable for any
      specification-constrained formatting evaluation.

\item To our knowledge, the first empirical comparison of four digit-generation algorithms
      (three fixed-width: \schubfach, \ryualg, \dragonbox; one
      multiprecision: \burgerdybvig) under \ecma/\rfcjcs constraints
      inside a canonicalization pipeline, showing that all three
      fixed-width algorithms perform within 1--8\% of each other while
      each achieves 1.4--1.8$\times$ speedup over \burgerdybvig on
      number-dense workloads ($\nu(D) \geq 40\%$) in Go API benchmarks;
      the advantage replicates at 2.3--3.0$\times$ in Rust.  The
      fixed-width equivalence is confirmatory (consistent with Adams
      and Champagne Gareau \& Lemire), but the pipeline context and
      Amdahl's Law characterization showing how isolation-level
      speedups attenuate at the system level are new.
      String-dominant and deeply-nested inputs show $<$5\% difference.

\item A cross-language validation design (4$\times$2 matrix of
      algorithm $\times$ language) that separates algorithmic effects from
      language/runtime effects, with Go cross-compilation determinism
      verified for x86-64 and arm64 via QEMU user-mode emulation.
      All evidence is released as a reproducibility artifact.
\end{enumerate}
