\section{Related Work}
\label{sec:related-work}

Extensive prior work has optimized float-to-string conversion in
isolation.  To our knowledge, this paper is the first to evaluate these algorithms
inside a specification-constrained canonicalization pipeline where
conformance is a binary gate.  We summarize the most relevant prior work, organized by
the problem dimension each addresses.

\subsection{Multiprecision Foundations}

Steele and White~\cite{steele1990how} established the foundational
framework for accurate float-to-string conversion.
Burger and Dybvig~\cite{burger1996printing} extended this to
shortest-representation printing using arbitrary-precision
rational arithmetic; their algorithm is our multiprecision baseline.
Gay's \texttt{dtoa.c}~\cite{gay1990dtoa} provides the foundational
practical implementation and remains widely used in language runtimes as
a reference or fallback.  All three works evaluate algorithms in isolation and
do not consider performance inside a pipeline where formatting policy is
externally constrained.

\subsection{Fixed-Width Approaches}

Loitsch~\cite{loitsch2010grisu} demonstrated that integer-only arithmetic
suffices for the common case with Grisu.  However, Grisu3 falls back to
arbitrary precision for $\sim$0.5\% of inputs, making conformance
guarantees harder to establish in specification-constrained contexts.

Adams~\cite{adams2018ryu} proposed Ry\=u, achieving full coverage with
fixed-width 128-bit arithmetic and lookup tables.  Adams provides the
most comprehensive prior performance comparison of fixed-width vs.\
multiprecision approaches (we refer the reader to Adams's literature
review rather than replicating it), but evaluates algorithms in isolation
on microbenchmarks without specification constraints.

Giulietti~\cite{giulietti2022schubfach} introduced the \schubfach
algorithm, which we implement and evaluate within the \rfcjcs pipeline.
Jeon~\cite{jeon2021dragonbox} extends and refines the \schubfach approach
with Dragonbox, improving table compression and branch structure.  Go~1.26
adopted Dragonbox for \texttt{strconv}~\cite{godragonboxcl}, providing
additional motivation for evaluating fixed-width approaches in
specification-constrained contexts.

\subsection{Comprehensive Experimental Review}

Champagne Gareau and Lemire~\cite{champagnegareau2026} provide a comprehensive experimental comparison of modern float-to-string
algorithms, evaluating Dragon4, Grisu3, Ry\=u, \schubfach, Dragonbox,
and \texttt{std::to\_chars} in C++.  Their work evaluates algorithms in
isolation on microbenchmarks.  The present study complements theirs by
embedding the same algorithms inside a specification-constrained
canonicalization pipeline where (a)~\ecma formatting policy is a hard
constraint, (b)~correctness against 286{,}362 oracle vectors is a
binary gate, and (c)~digit generation is only one component of the
total cost.

The complementary direction, fast string-to-float parsing, is addressed
by Lemire~\cite{lemire2021number}, whose Eisel-Lemire algorithm achieves
gigabyte-per-second parsing throughput.  While outside our scope, this
work shares the fixed-width arithmetic philosophy that motivates
\schubfach and Dragonbox.

\subsection{Deterministic Serialization Standards}

Beyond JSON, deterministic encoding is required by other serialization
formats.  CBOR (RFC~8949)~\cite{rfc8949} specifies deterministic encoding
rules (\S4.2) for binary data, addressing analogous concerns about
canonical representation.  W3C RDF Dataset
Canonicalization~\cite{w3c2024rdf} addresses graph-level canonicalization
(structural equivalence of RDF datasets), a structurally different
problem from the byte-level serialization determinism required by
\rfcjcs.  Our conformance-gated methodology could be adapted to other
deterministic serialization benchmarks.

\subsection{JSON Canonicalization}

\rfcjcs~\cite{rfc8785} is the primary standard.  Prior to RFC~8785,
ad-hoc canonicalization schemes existed in various ecosystems (e.g.,
JSON-LD canonicalization, OLPC canonical JSON).  None specify the
floating-point formatting requirement with the precision of
\ecma~\cite{ecma262}.

Prior to this work, conformance verification against \ecma oracle
vectors revealed that existing \rfcjcs implementations couple number
formatting to language runtime internals without implementing the \ecma
formatting policy directly.  The standard pattern (delegating to the
runtime's float-to-string function and post-processing the result) is
fragile: Go's \texttt{strconv.FormatFloat}, the dependency in a widely
used Go \rfcjcs implementation, has undergone three
algorithm changes (Grisu3 through Go~1.13, Ry\=u from Go~1.14,
Dragonbox from Go~1.26) and had a confirmed rounding bug (Go
issue~\#29491, filed by the \rfcjcs author, where \texttt{FormatFloat}
returned wrong digits for specific values on certain platforms).  This
ecosystem gap motivated both the custom implementation and the
conformance-gated methodology presented here.

\subsection{Benchmark Methodology}

Our conformance-before-performance approach draws on the principle that
correctness is a prerequisite for meaningful performance comparison.
Georges et al.~\cite{georges2007statistically} established rigorous
statistical methods for Java performance evaluation, advocating
confidence intervals and effect sizes over single-point estimates.
Mytkowicz et al.~\cite{mytkowicz2009producing} demonstrated that
measurement bias (link order, environment size) can reverse performance
conclusions, motivating our randomized task ordering and environment
documentation.  Kalibera and Jones~\cite{kalibera2013rigorous} formalized
the analysis of multi-level experimental designs in benchmarking.  Our
use of permutation tests and BCa bootstrap confidence intervals follows
this modern statistical practice for performance evaluation.
