\section{Results}
\label{sec:results}

% Tables will be generated from benchmark data by scripts/generate-paper-tables.go
% and placed in paper/tables/

\subsection{Conformance}

All eight implementations (Go and Rust $\times$ \schubfach, \ryualg,
\dragonbox, and \burgerdybvig) pass the full conformance matrix: 776 total
cases (97~structural tests $\times$ 8~implementations, 0~failures).
Output SHA-256 digests are byte-identical across all implementations for
valid inputs.

\subsection{API Benchmarks}

API benchmarks isolate the algorithmic cost of the canonicalization
function, removing process startup and I/O overhead.  These measurements
most directly reflect the digit-generation performance difference.

Table~\ref{tab:api-summary} summarizes the key API results for the
within-language algorithm comparisons (canonicalize mode).
Figure~\ref{fig:api-speedup} visualizes the speedup across all shared
workloads for both language pairs.  Full results
for all pairs, modes, and workloads appear in
Appendix~\ref{sec:appendix-tables}.

\input{tables/api-summary}

Table~\ref{tab:api-stats-summary} presents the within-language API
comparisons with BH-corrected $p$-values, confidence intervals, and
effect sizes.

\input{tables/api-stats-summary}

\input{figures/api-speedup-comparison}

\subsection{Fixed-Width Algorithm Equivalence}

A central finding is that the three fixed-width algorithms (\schubfach,
\ryualg, \dragonbox) perform within 1--8\% of each other across both
languages and all workloads.  In Go API benchmarks, within-fixed-width
comparisons (go:schubfach-vs-ryu, go:schubfach-vs-dragonbox,
go:ryu-vs-dragonbox) show speedups of ${\leq}1.02\times$ with few
BH-significant differences.  In Rust, the same pattern holds, with
occasional small differences (${\leq}1.13\times$) reflecting
compilation artifacts rather than algorithmic cost.

This equivalence confirms that the dominant performance axis is
\emph{fixed-width vs.\ multiprecision}, not the choice among
fixed-width algorithms.

\subsection{CLI Benchmarks}

CLI benchmarks capture end-to-end pipeline cost including process
startup, argument parsing, file I/O, and output serialization.  Estimated
process startup overhead (approximately 1.3\,ms for Go, 0.5\,ms for Rust)
dominates for smaller workloads, attenuating the algorithmic signal
observed in API benchmarks.

Table~\ref{tab:cli-summary} summarizes the key CLI results.
Figure~\ref{fig:cli-significance} shows the end-to-end speedup with
BH-significance markers.  Full results appear in
Appendix~\ref{sec:appendix-tables}.

\input{tables/cli-summary}

\input{figures/cli-significance}

\subsection{Statistical Comparisons}

Within-language CLI comparisons (canonicalize mode) with BH-corrected
$p$-values, confidence intervals, and effect sizes are reported in
Table~\ref{tab:stats-summary} (Appendix~\ref{sec:appendix-tables}).
The patterns mirror the API results: fixed-width algorithms outperform
\burgerdybvig on number-dense workloads, with smaller effect sizes due
to process startup dilution.  The full 1{,}644-comparison statistical
report appears in the supplementary materials.

\subsection{Cross-Language Validation}

All eight implementations pass the full conformance matrix (776 total
cases: 97~structural tests $\times$ 8~implementations, 0~failures) and
produce byte-identical output on all oracle vectors.  2{,}000
differential fuzz cases yield zero cross-implementation divergences.

The statistical report includes 16 comparison classes organized into
three groups:
\begin{itemize}
\item 6~within-Go algorithm pairs and 6~within-Rust algorithm pairs
      (testing algorithmic effects within each language),
\item 4~cross-language same-algorithm pairs (testing language/runtime
      effects while controlling for algorithm).
\end{itemize}

In CLI end-to-end measurement, Rust implementations are approximately
2--3$\times$ faster than their Go counterparts, primarily due to lower
process startup overhead (Rust native binary vs.\ Go runtime
initialization).  In API benchmarks, the within-Rust algorithmic ordering
replicates the within-Go ordering: all three fixed-width algorithms
outperform \burgerdybvig on number-dense workloads; this replication
of within-language ordering provides evidence that the fixed-width
advantage is not solely attributable to language-specific runtime
characteristics.

This decomposition separates algorithmic effects from language/runtime
effects and directly addresses language-specific performance threats.

\subsection{Differential Fuzzing}

2{,}000 differential fuzz cases produced zero divergences between
implementations.
