\section{Threats to Validity}
\label{sec:threats}

\subsection{Internal Validity}

\textbf{Warmup validation.}
CLI benchmarks discard 3~warmup runs.  Rust API benchmarks discard
3~warmup iterations per workload to stabilize instruction caches,
branch predictors, and lazily-initialized data structures (specifically,
the Rust \burgerdybvig implementation's \texttt{OnceLock}-based
power-of-ten cache).  Go API benchmarks benefit from \texttt{testing.B}'s
built-in warmup.  Per-iteration timing data
(Table~\ref{tab:per-iteration-outliers}) confirms low coefficient of
variation ($<$5\%) across measured iterations, indicating stable
steady-state behavior after warmup.  Figure~\ref{fig:warmup-validation}
shows 20-iteration warmup validation data for a representative workload.

\input{figures/warmup-validation}

\textbf{Sequential run autocorrelation.}
CLI benchmarks run sequentially on the same machine; consecutive runs
share CPU thermal state, cache warmth, and potentially TDP-limited
frequency.  The randomized task order (fixed seed) mitigates systematic
ordering effects, but thermal coupling between adjacent runs creates
weak positive autocorrelation that may slightly inflate the effective
sample size.  Permutation tests are robust to mild non-stationarity but
assume exchangeability under the null hypothesis; this assumption is
approximately but not exactly satisfied.

\textbf{CPU core affinity.}
CLI benchmarks ran without CPU pinning on heterogeneous P-core/E-core
topology (8~P-cores at 3.6--5.0\,GHz, 4~E-cores at 2.7--3.8\,GHz).
Process migration may inflate variance for small workloads near the
noise floor; an E-core placement could slow a single run by 30--40\%.
The high MDEs observed for small Go CLI workloads (14--24\%) are
consistent with this variance source.  API benchmarks (in-process) are
less susceptible.

\textbf{API measurement granularity.}
Go API benchmarks (\texttt{testing.B}) auto-calibrate the number of
inner loop iterations per sample, producing averaged low-variance
samples.  Rust API benchmarks time individual function calls ($N{=}1$
per sample), producing inherently higher-variance samples.  This
asymmetry affects cross-language comparisons but not within-language
comparisons.  We note this caveat in \S\ref{sec:discussion} and
emphasize within-language results for algorithmic conclusions.

\textbf{sync.Pool mitigation.}
The Go \burgerdybvig implementation uses \texttt{sync.Pool} to reuse
big-integer buffers, reducing allocation cost.  Without pooling, the
performance gap would be wider.  This reflects realistic production usage
but may understate the theoretical algorithmic advantage of fixed-width
approaches.

\subsection{Construct Validity}

\textbf{Allocation instrumentation.}
Rust allocation data is measured via a counting global allocator
(\texttt{std::alloc::GlobalAlloc}) that intercepts all heap allocations
and reports per-operation byte and count totals.  This includes
allocations from the benchmark infrastructure itself (e.g., output buffer
construction), not solely the digit-generation path.  Go's
\texttt{testing.B} has the same characteristic.  Within-language
$\Delta$allocs comparisons are valid because both implementations share
the same infrastructure overhead.

\textbf{Compiler optimization asymmetry.}
Rust builds use \texttt{opt-level=3}, link-time optimization (LTO), and
\texttt{codegen-units=1}, enabling aggressive cross-crate inlining and
whole-program optimization.  Go's compiler does not support LTO and has
a less aggressive inlining heuristic.  The observed cross-language
speedups (2--4$\times$ in API benchmarks) may partly reflect differences
in compiler optimization quality rather than intrinsic language
characteristics.  Within-language comparisons are unaffected since both
implementations in each pair share the same compiler pipeline.

\textbf{Memory allocator differences.}
Go uses a tcmalloc-derived allocator integrated with its garbage
collector.  Rust uses the system allocator (typically glibc malloc on
Linux).  For the multiprecision (\burgerdybvig) path, which performs
heap allocations for big-integer arithmetic, allocator performance
differences could affect cross-language comparisons.  Within-language
comparisons are unaffected since both implementations share the same
allocator.

\textbf{Multiprecision implementation quality.}
The Rust \burgerdybvig implementation uses the \texttt{num-bigint}
crate (v0.4), which employs Karatsuba multiplication for large
operands, comparable in optimization level to Go's
\texttt{math/big.Int} with its assembly-accelerated x86-64 inner
loops.  Both languages therefore use production-grade multiprecision
libraries, making within-Rust and within-Go speedup ratios directly
comparable without an implementation-quality confound.

\textbf{Go toolchain version.}
Benchmarks use Go~1.25.6.  Since neither implementation uses
\texttt{strconv.FormatFloat}, Go~1.26's adoption of Dragonbox for
\texttt{strconv} does not affect the digit-generation paths under test.
The Go compiler version affects only runtime infrastructure (allocator,
scheduler, garbage collector), which is shared identically between
both implementations within each pair.

\textbf{Oracle SHA-256 circularity.}
SHA-256 checksums verify vector integrity but cannot verify oracle
correctness.  This threat is mitigated by cross-engine validation:
all 286{,}362 vectors produce identical output under V8~12.9.202.28
(Node.js v23.3.0), SpiderMonkey 115.10.0, and JavaScriptCore
(WebKitGTK~2.50.4), confirming that the oracle reflects ECMA-262
\texttt{Number::toString} semantics rather than a single engine's
idiosyncratic behavior.

\subsection{External Validity}

\textbf{Language specificity (mitigated).}
This threat is substantially mitigated by the $4 \times 2$
cross-language comparison matrix: all four algorithms are implemented in
Go and Rust, and the fixed-width performance ordering holds in both
languages.  Within-Rust
comparisons replicate the within-Go algorithmic advantage, confirming
that the result is not an artifact of Go's garbage collector or
allocation subsystem.  Residual external-validity risk remains for
additional runtimes (e.g., JVM/generational-GC and unmanaged C
implementations) that are not yet benchmarked.

\textbf{Single-architecture performance data.}
Performance benchmarks run on a single x86-64 machine.
Cross-compilation \emph{determinism} is verified by building arm64
binaries and executing them under QEMU user-mode emulation on the x86-64
host.  Because QEMU user-mode translates system calls but executes
floating-point operations on the host CPU, this verifies Go compiler
determinism across target architectures, not native arm64 hardware
behavior.  Arm64 \emph{performance} benchmarks are deferred due to lack
of bare-metal hardware.  Future work should include bare-metal arm64
benchmarks on Graviton, Apple Silicon, or Ampere hardware.

\textbf{JavaScript engine oracle (mitigated).}
Oracle vectors are generated from V8 (via Node.js v23.3.0,
V8~12.9.202.28) and independently validated against SpiderMonkey
(JavaScript-C115.10.0) and JavaScriptCore (WebKitGTK~2.50.4).  All
three engines, representing the three major independent implementations
of ECMA-262 \texttt{Number::toString}, produce identical output for
all 286{,}362 vectors with zero divergences.  Combined with dense
boundary coverage (all subnormals up to $2^{12}$, all binary64 powers
of two, subnormal/normal boundary neighborhoods, \rfcjcs Appendix~B
examples with neighbors, and pseudorandom fill), this provides
evidence that the oracle reflects the normative specification rather
than engine-specific behavior.  Cross-version stability within each
engine family is expected from the normative spec but not independently
verified.

\textbf{Workload representativeness.}
The benchmark suite includes both synthetic workloads designed to bracket
the performance space between number-heavy and string-dominant extremes,
and five production-representative workloads drawn from public
specifications (GeoJSON, Kubernetes pod specs, npm package metadata,
CloudEvents envelopes, and OpenAPI schema fragments).  The production
workloads span number densities from ${\approx}5\%$ (package metadata)
to ${\approx}70\%$ (GeoJSON coordinates) and fall on the same
density--speedup trend line as the synthetic workloads, confirming that
the synthetic results generalize to realistic payloads.

\textbf{Endianness.}
Both x86-64 and arm64 targets are little-endian; big-endian behavior is
untested.  While the canonicalization algorithms do not depend on byte
order (they operate on IEEE~754 binary64 values via language-level
primitives), edge cases in binary representation parsing could
theoretically differ on big-endian platforms.
