\documentclass[12pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{xcolor}

\signature{Mark Lenhardt\\Lattice Substrate}
\address{Lattice Substrate}

\begin{document}

\begin{letter}{%
The Editors\\
\textit{Software: Practice and Experience}\\
John Wiley \& Sons}

\opening{Dear Editors,}

Please consider the enclosed manuscript, ``Fixed-Width vs.\
Multiprecision Digit Generation for JSON Canonicalization: A
Four-Algorithm Empirical Study,'' for publication in \textit{Software:
Practice and Experience}.

The paper makes three contributions to the practice of
specification-constrained performance evaluation:

\begin{enumerate}
\item \textbf{Conformance-gated benchmarking}, a methodology that
      treats specification correctness as a binary prerequisite for
      performance measurement.  The methodology was developed from a
      practical discovery: existing JSON Canonicalization Scheme
      (RFC~8785) implementations delegate number formatting to language
      runtimes that silently violate the required ECMA-262
      \texttt{Number::toString} policy, making speed comparisons
      meaningless without a conformance gate.

\item \textbf{A pipeline-level empirical study} comparing four
      digit-generation algorithms (Schubfach, Ry\=u, Dragonbox, and
      Burger--Dybvig) inside an RFC~8785 canonicalization pipeline,
      with cross-language validation in Go and Rust (8~implementations,
      1,644~statistical comparisons).  This characterizes how Amdahl's
      Law attenuates isolation-level algorithm speedups at the system
      level---a perspective absent from prior isolation benchmarks.

\item \textbf{A cross-language validation design} (4$\times$2 matrix
      of algorithm $\times$ language) that separates algorithmic effects
      from language/runtime characteristics.
\end{enumerate}

The work is practice-oriented: the methodology emerged from building a
production canonicalization primitive, and the findings directly inform
algorithm selection for implementors.  The paper engages with Champagne
Gareau and Lemire's recent float-to-string survey published in
\textit{SP\&E} (January~2026), extending their isolation-level results
to a pipeline context with conformance gating.

All source code (8~implementations), 286,362 oracle vectors, raw
benchmark data, and the statistical analysis pipeline are archived at
Zenodo (DOI: \href{https://doi.org/10.5281/zenodo.18890835}{10.5281/zenodo.18890835})
and available on GitHub at
\href{https://github.com/lattice-substrate/jcs-bench-lab}{github.com/lattice-substrate/jcs-bench-lab}.

This manuscript has not been submitted to or published in any other
journal or conference proceedings, and the sole author has approved
the submission.

\closing{Sincerely,}

\end{letter}
\end{document}
