% NOTE: For final SP&E submission, reformat using Wiley's LaTeX template
% (wiley-article class). This draft uses standard article class.
\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{caption}
\pgfplotsset{compat=1.18}

\newcommand{\schubfach}{Schubfach\xspace}
\newcommand{\burgerdybvig}{Burger--Dybvig\xspace}
\newcommand{\ryualg}{Ry\=u\xspace}
\newcommand{\dragonbox}{Dragonbox\xspace}
\newcommand{\ecma}{ECMA-262\xspace}
\newcommand{\rfcjcs}{RFC~8785\xspace}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}

\title{Fixed-Width vs.\ Multiprecision Digit Generation\\for JSON Canonicalization: A Four-Algorithm Empirical Study}

\author{Mark Lenhardt}
\affil{Lattice Substrate}
\date{}

\maketitle

\begin{abstract}
We introduce \emph{conformance-gated benchmarking}, a methodology
that treats specification correctness as a binary prerequisite for
performance measurement, and characterize how Amdahl's Law attenuates
isolation-level algorithm speedups inside a real formatting pipeline.
The methodology was developed from a practical discovery: existing
JSON Canonicalization Scheme (RFC~8785) implementations delegate
number formatting to language runtimes that silently violate the
required ECMA-262 \texttt{Number::toString} policy, so benchmarking
them without first establishing conformance produces meaningless
speed comparisons.  Each candidate is gated on an oracle vector suite,
noise floors and minimum detectable effect sizes are documented for
every comparison, and false-discovery rate is controlled across the
full comparison matrix.  The methodology is general, applicable
wherever correctness is a hard prerequisite.

We apply it to a case study comparing four digit-generation algorithms
(three fixed-width: \schubfach, \ryualg, \dragonbox; one
multiprecision: \burgerdybvig) inside an RFC~8785 canonicalization
pipeline, with cross-language validation in Go and Rust
(8~implementations, 1{,}644 statistical comparisons).  Conformance is
verified against 286{,}362 oracle vectors generated from V8 and
validated against SpiderMonkey and JavaScriptCore.  The three
fixed-width algorithms perform within 1--8\% of each other across
both languages.  Each is 1.4--1.8$\times$ faster than \burgerdybvig
on number-dense workloads ($\nu(D) \geq 40\%$) in Go API benchmarks,
replicating at 2.3--3.0$\times$ in Rust (excluding two pure-numeric
outliers discussed in \S\ref{sec:discussion}).  Profiling shows digit
generation accounts for 15--40\% of total pipeline cost; the observed
ratios match the Amdahl's Law prediction at these pipeline fractions,
demonstrating that fixed-width algorithms have largely exhausted the
available improvement from optimizing digit generation alone.  On
string-dominant and deeply-nested inputs the difference narrows to
$<$5\%.  All eight implementations produce byte-identical output; Go
binaries are additionally verified under cross-compilation for x86-64
and arm64.  We release the benchmark infrastructure, oracle vectors,
and all raw data as a reproducibility artifact.
\end{abstract}

\noindent\textbf{Keywords:} JSON canonicalization, floating-point formatting, Schubfach,
Ry\=u, Dragonbox, Burger--Dybvig, RFC~8785, ECMA-262, benchmark methodology, cross-language validation

\input{sections/introduction}
\input{sections/background}
\input{sections/related-work}
\input{sections/algorithms}
\input{sections/implementation}
\input{sections/methodology}
\input{sections/results}
\input{sections/discussion}
\input{sections/threats}
\input{sections/conclusion}

\section*{Data Availability Statement}

All data, software, and oracle vectors are freely available online.
The benchmark infrastructure, all eight implementations, workload manifests,
and oracle generation scripts are available at the accompanying
repository.\footnote{\url{https://github.com/lattice-substrate/jcs-bench-lab}.
Archived at Zenodo: \url{https://doi.org/10.5281/zenodo.18890835}.}  Complete benchmark tables (full API benchmarks,
full CLI benchmarks, and all 1{,}644 statistical comparisons across
16~pairs) are provided in the supplementary materials.  Oracle vector
checksums (SHA-256) are documented in \texttt{PROVENANCE.json} files
accompanying each implementation.

\section*{Author Contributions}

Following the CRediT (Contributor Roles Taxonomy) format:
\textit{Mark Lenhardt:} conceptualization; methodology;
software; validation; formal analysis; investigation; data curation;
writing (original draft); writing (review and editing); visualization.

\bibliographystyle{plain}
\bibliography{references}

\clearpage
\appendix
\input{sections/appendix-tables}
\input{sections/appendix-reproducibility}

\end{document}
