% Supplementary Materials for:
% Fixed-Width vs. Multiprecision Digit Generation for JSON Canonicalization:
% A Four-Algorithm Empirical Study
%
% Mark Lenhardt, Lattice Substrate
\documentclass[10pt]{article}

\usepackage[margin=0.75in]{geometry}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{fancyhdr}

\newcommand{\schubfach}{Schubfach\xspace}
\newcommand{\burgerdybvig}{Burger--Dybvig\xspace}
\newcommand{\ryualg}{Ry\=u\xspace}
\newcommand{\dragonbox}{Dragonbox\xspace}
\newcommand{\ecma}{ECMA-262\xspace}
\newcommand{\rfcjcs}{RFC~8785\xspace}

% Provide labels referenced by auto-generated tables that are defined in the main paper
\newcommand{\fakesec}[1]{\label{#1}}
\AtBeginDocument{%
  \fakesec{sec:discussion}%
  \fakesec{sec:appendix-tables}%
  \fakesec{sec:methodology}%
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Supplementary Materials --- Lenhardt (2026)}
\fancyhead[R]{\small \thepage}
\renewcommand{\headrulewidth}{0.4pt}

\begin{document}

\begin{center}
{\Large\bfseries Supplementary Materials}\\[6pt]
{\large Fixed-Width vs.\ Multiprecision Digit Generation\\
for JSON Canonicalization: A Four-Algorithm Empirical Study}\\[6pt]
Mark Lenhardt\\
Lattice Substrate\\[6pt]
\today
\end{center}

\vspace{1em}
\noindent
This document contains the complete benchmark tables and statistical
comparisons referenced as ``supplementary materials'' in the main
paper.  All data are auto-generated from raw benchmark output by
\texttt{scripts/generate-paper-tables.go} and are reproducible from
the archived artifact (DOI:
\href{https://doi.org/10.5281/zenodo.18890835}{10.5281/zenodo.18890835}).

\tableofcontents

\clearpage

% ─── Section 1: Full API Benchmarks ───────────────────────────────────────────

\section{Full API Benchmark Results}
\label{sec:supp-api}

Table~\ref{tab:api-benchmarks} reports the complete API benchmark
results for all 16~comparison pairs across all workloads and modes
(canonicalize and verify).  Each row reports the mean of $n{=}10$
iterations.  Allocation columns ($\Delta$allocs/op, $\Delta$B/op)
report the difference between the two implementations in each pair;
negative values indicate the first implementation allocates less.

For Rust implementations, allocation data is measured via a counting
global allocator (\texttt{std::alloc::GlobalAlloc}) that intercepts
all heap allocations.

\begin{landscape}
\input{tables/api-benchmarks}
\end{landscape}

\clearpage

% ─── Section 2: Full CLI Benchmarks ──────────────────────────────────────────

\section{Full CLI Benchmark Results}
\label{sec:supp-cli}

Table~\ref{tab:cli-benchmarks} reports the complete CLI benchmark
results for all 16~comparison pairs across all workloads and modes.
Each row reports the mean wall time of $n{=}15$ repetitions (3~warmup
discarded).  Task order was randomized with a fixed seed.

\begin{landscape}
\input{tables/cli-benchmarks}
\end{landscape}

\clearpage

% ─── Section 3: Full Statistical Comparisons ─────────────────────────────────

\section{Full Statistical Comparisons}
\label{sec:supp-stats}

Table~\ref{tab:stats-comparisons} reports all 1{,}644 statistical
comparisons across 16~pairs, 3~tracks (API, CLI end-to-end, CLI
algorithmic), 2~modes (canonicalize, verify), and all workloads.
Each comparison includes:

\begin{itemize}
\item \textbf{Speedup:} ratio of slower to faster mean.
\item \textbf{CI$_{95}$:} BCa bootstrap 95\% confidence interval on
      the speedup ratio (5{,}000 resamples).
\item \textbf{$p$:} raw permutation test $p$-value (5{,}000 resamples).
\item \textbf{$p_\text{adj}$:} Benjamini--Hochberg adjusted $p$-value
      controlling false discovery rate at $q = 0.05$ across all
      1{,}644 comparisons.
\item \textbf{$d$:} Cohen's $d$ effect size
      ($(\bar{x}_A - \bar{x}_B)/s_p$; negative = implementation A
      faster).
\item \textbf{Noise:} standard deviation of the faster implementation,
      quantifying measurement noise in the same units as the metric.
\item \textbf{MDE:} minimum detectable effect at 80\% power, expressed
      as a percentage of the faster implementation's mean.
\end{itemize}

\noindent
$p$-values at the resolution floor of $1/(R+1) = 0.0002$ (where
$R = 5{,}000$ resamples) should be interpreted as $p < 0.0002$.

\begin{landscape}
\input{tables/stats-comparisons}
\end{landscape}

\end{document}
